Supporting Information

### 2018-04-19 SupportingInformation.Rmd
modified

Carbon and nitrogen in soil at many locations across CONUS (Terry to augment this description). The raw data are shown in Figure \ref{fig_soilCNraw}. The copula is shown in Figure \ref{fig_soilCNcop}.

<!--The raw data look like this:-->
```{r load_data_Loecke, echo=F, results='hide'}
#fig.cap="Plot of the raw soil C and N data.\\label{fig_soilCNraw}", fig.height=3, fig.width=3}
d<-readRDS("Data/RaCA_soilorganicC_soiltotalN_stocks100cm.RDS")
d<-d[,c("SOCstock100","TSNstock100")]
pdf("./Results/SoilCNraw.pdf")
plot(d[,1],d[,2],type="p",col="red",xlab="Soil C",ylab="Soil N",cex.lab=1.5,cex.axis=1.5)
dev.off()
```

<!--The copula looks like this:-->
```{r rankandplot_Loecke, echo=F, results='hide'}
#fig.cap="Copula plot of soil C and N data. Obtained by applying ranks separately to the C and N data, and then dividing the ranks by $n+1$ where $n$ is the number of sampling locations (see Sec. \ref{Model_selection}).\\label{fig_soilCNcop}",  fig.height=5, fig.width=5}
source("getcopula.R")
#convert to a copula and plot
pdf("./Results/SoilCNcop.pdf")
v_CN<-getcopula(d=d,rankon=T,ploton=T) 
dev.off()
```

![Plot of the raw soil C and N data.\label{fig_soilCNraw}](./Results/SoilCNraw.pdf){ width=30%, height=30% }

![Copula plot of soil C and N data. Obtained by applying ranks separately to the C and N data, and then dividing the ranks by $n+1$ where $n$ is the number of sampling locations (see Sec. \ref{Model_selection}). \label{fig_soilCNcop}](./Results/SoilCNcop.pdf){ width=30%, height=30% }

## Phylogenetic data  \label{Phylo_data}
```{r calculation_phylo_data, echo=F, results="hide"}
source("phylo_dat_fn.R")
phylo_dat<-read.csv("./Data/Anolis/contrasts.csv",header = T)

# This plot I create on temporary basis to see the copula structure between various column of the raw data
# After I know which variables should I look into, I will modify this chunk and add model selection and NPA on that
pdf("./Results/temp_phylo_plot.pdf",width=20, height=20)
op<-par(mfrow=c(13,13))
for(i in c(1:13)){
  for(j in c(1:13)){
    v_phylo<-phylo_dat_fn(data=phylo_dat,i=i,j=j,ploton = F)
    plot(v_phylo[,1],v_phylo[,2],col="red")
    #mtext("(i,j)=(",i,", ",j,")")
    mtext(paste0(" [",i,",",j,"] ", ","," n=",dim(v_phylo)[1]),side = 3,col="blue")
  } 
}
par(op)
dev.off()
```

@@

We also tested statistically for asymmetry of tail dependence, which seems visible in Figure \ref{fig_soilCNcop}. 
Results (Table \ref{tab_soilCN_asym}) show that upper-tail dependence was, indeed, significantly stronger than lower-tail dependence in these data, answering the third part of question 1.

@@

 atdres<-data.frame(Statistic=c("$\\cor_{0,0.1}-\\cor_{0.9,1}$",
                                "$\\Ps_{0,0.1}-\\Ps_{0.9,1}$",
                                "$\\Dsq_{0.9,1}-\\Dsq_{0,0.1}$"),
 #atdres<-data.frame(Statistic=c("$\\cor_{first.slice}-\\cor_{last.slice}$",
 #                              "$\\Ps_{first.slice}-\\Ps_{last.slice}$",
 #                               "$\\Dsq_{first.slice}-\\Dsq_{last.slice}$"),
                   Kendall=rep("",3),
                   Spearman=rep("",3),stringsAsFactors = F)

@@

```{r read_plot_BMR_data, echo=FALSE, results="hide"}
set.seed(seed)
source("phylo_dat_fn.R")
#---------------------------------------------------------
BMR_dat<-read.csv("./Data/BMR/bmr-contrasts.csv",header = T)
#class(BMR_dat)
v_BMR<-phylo_dat_fn(data=BMR_dat,i=4,j=3,ploton = F)

pdf("./Results/BMR_results/BMR_vs_lnmass.pdf")
op<-par(mar=c(5,5,5,5))
plot(v_BMR[,1],v_BMR[,2],col="red",xlab="ln(mass)",ylab="BMR",cex.lab=2,cex.axis=2) # copula plot btw ln(Mass) vs. BMR
par(op)
dev.off()
```

![Copula plot for BMR dataset \label{fig_BMR_cop}](./Results/BMR_results/BMR_vs_lnmass.pdf){ width=30%, height=30% }

@@

@@

@@

```{r stat_BMR_graphic_approach, echo=F, results="hide", cache=T, cache.extra=list(seed,v_BMR,mtime("bivfunctionplot.R"),mtime("CopulaFunctions_flexible.R"),mtime("copsurrog2d.R"))}
source("bivfunctionplot.R")
set.seed(seed) 
# Now do NPA for v_BMR
# If "F"/"SBB8" is the best fit : then it should be symmetric tail dep (0) : Then still I need to do NPA ?
# May be : because it will also tell me how they are different from "N" copula in different sections?
source("bivfunctionplot.R")
stat_BMR_ga<-bivfunctionplot(v=v_BMR,resloc="./Results/BMR_results/",nametag="BMR",numbin = 10)
```

![Nonparametric tests for tail dependence and other deviations from a normal copula for BMR data. As described in Methods, each statistic was computed on the data using the ranges $0-0.1$, $0.1-0.2$, ..., $0.9-1$ (red lines), and was also computed using the same ranges for 1000 normal-copula surrogate datasets (dashed lines show 0.025 and 0.975 quantiles of these results and dotted lines show 0.005 and 0.995 quantiles). \label{fig_BMR_nonparam}](./Results/BMR_results/BMR_bivfunctionplot.pdf)

```{r tab_BMR_asym, echo=F, results="markup"}
tab_BMR_asym<-make_table_stat_fn(data=stat_BMR_ga)
knitr::kable(tab_BMR_asym,
             format="pandoc",
             caption = "Results for tests of whether asymmetry of tail dependence was significant compared to a normal-copula null hypothesis (best fit is here Frank with symmetric tail dep)? \\label{tab_BMR_asym}",
             booktabs=T)
```

>>>>

Carbon and nitrogen in soil at many locations across CONUS (Terry to augment this description). The raw data are shown in Figure \ref{fig_raw_cop_soilCN}(A). The copula is shown in Figure \ref{fig_raw_cop_soilCN}(B).

<!--The raw data look like this:-->
```{r load_data_Loecke, echo=F, results='hide'}
#fig.cap="Plot of the raw soil C and N data.\\label{fig_raw_cop_soilCN}(A)", fig.height=3, fig.width=3}
d<-readRDS("Data/RaCA_soilorganicC_soiltotalN_stocks100cm.RDS")
d<-d[,c("SOCstock100","TSNstock100")]
pdf("./Results/SoilCNraw.pdf")
op<-par(mar=c(5.1, 4.1, 0.2, 2.1))
plot(d[,1],d[,2],type="p",col="red",xlab="Soil C",ylab="Soil N",cex.lab=1.5,cex.axis=1.5)
par(op)
dev.off()
```

<!--The copula looks like this:-->
```{r rankandplot_Loecke, echo=F, results='hide'}
source("getcopula.R")
#convert to a copula and plot
pdf("./Results/SoilCNcop.pdf")
v_CN<-getcopula(d=d,rankon=T,ploton=T) 
dev.off()
```

\begin{figure}[!h]
\textbf{ \hspace{5.5 cm} (A) \hspace{7 cm} (B)  } \\
\begin{center}
\includegraphics[width=3 in]{./Results/SoilCNraw.pdf}
\includegraphics[width=3 in]{./Results/SoilCNcop.pdf}
\caption{\textbf{(A)} Plot of the raw soil C and N data, \textbf{(B)} Copula plot of soil C and N data. Obtained by applying ranks separately to the C and N data, and then dividing the ranks by $n+1$ where $n = `r nrow(d)` $ is the number of sampling locations (see Sec. \ref{Model_selection}).\label{fig_raw_cop_soilCN}}
\end{center}
\end{figure}

## BMR data  \label{BMR_data}

```{r read_plot_BMR_data, echo=FALSE, results="hide"}
set.seed(seed)
source("phylo_dat_fn.R")
#---------------------------------------------------------
BMR_dat<-read.csv("./Data/BMR/bmr-contrasts.csv",header = T)
#class(BMR_dat)

# Generating raw data plot btw ln(mass) and BMR from bmr-contrasts.csv
pdf("./Results/BMR_results/rawdataplot_BMR_vs_lnmass.pdf")
op<-par(mar=c(5.1, 4.1, 0.2, 2.1))
plot(BMR_dat$ln.mass,BMR_dat$BMR,type="p",col="red",xlab="ln(mass)",ylab="BMR",
     cex.lab=1.5,cex.axis=1.5)
par(op)
dev.off()
#--------------------------------------------------------------------------------------------
# Generating raw data plot btw ln(mass) and BMR from bmr-contrasts.csv
pdf("./Results/BMR_results/copula_BMR_vs_lnmass.pdf")
v_BMR<-phylo_dat_fn(data=BMR_dat,i=4,j=3,ploton = T) 
#plot(v_BMR[,1],v_BMR[,2],type="p",col="red",xlab="ln(mass)",ylab="BMR",
#     cex.lab=1.5,cex.axis=1.5)
dev.off()
```

\begin{figure}[!h]
\textbf{ \hspace{5.5 cm} (A) \hspace{7 cm} (B)  } \\
\begin{center}
\includegraphics[width=3 in]{./Results/BMR_results/rawdataplot_BMR_vs_lnmass.pdf}
\includegraphics[width=3 in]{./Results/BMR_results/copula_BMR_vs_lnmass.pdf}
\caption{\textbf{(A)} Plot of the raw BMR vs. ln(mass) data \textbf{(B)} Copula plot for BMR dataset. Obtained by applying ranks separately to each variable, and then dividing the ranks by $n+1$ where $n=`r nrow(v_BMR)`$ is the number of contrasts (see Sec. \ref{Model_selection}).\label{fig_BMR_raw_cop}}
\end{center}
\end{figure}

## Phylogenetic data  \label{Phylo_data}

```{r calculation_phylo_data, echo=F, results="hide"}
source("phylo_dat_fn.R")
phylo_dat<-read.csv("./Data/Anolis/contrasts.csv",header = T)
#class(phylo_dat) #it's a data.frame
#----------------------------------------------------------------------------------------------------
# Generating all copula plots between different pathlength and traits 
pdf("./Results/Anolis_results/contrasts_csv_results/contrasts_csv_pathlength_traits_cop_plot.pdf",width=20, height = 20)
op<-par(mfrow=c(4,4))
i<-1
  for(j in c(1:13)){
    v_phylo<-phylo_dat_fn(data=phylo_dat,i=i,j=j,ploton = T)
    #plot(v_phylo[,1],v_phylo[,2],col="green4")
    #mtext("(i,j)=(",i,", ",j,")")
    mtext(paste0(" [",i,",",j,"] ", ","," n=",dim(v_phylo)[1]),side = 3,col="blue")
  } 
par(op)
dev.off()
#------------------------------------------------------------------------------------------------------------
# Generating all 12 by 12 copula plots between different traits
pdf("./Results/Anolis_results/contrasts_csv_results/contrasts_csv_btw_traits_cop_plot.pdf",width=20, height = 20)
op<-par(mfrow=c(12,12))
for(i in c(2:13)){
  for(j in c(2:13)){
    v_phylo<-phylo_dat_fn(data=phylo_dat,i=i,j=j,ploton = T)
    mtext(paste0(" [",i,",",j,"] ", ","," n=",dim(v_phylo)[1]),side = 3,col="blue")
  } 
}
par(op)
dev.off()
#----------------------------------------------------------------------------------------
# Generating raw data plot btw AVG.humerus and AVG.radius from contrasts.csv
pdf("./Results/Anolis_results/contrasts_csv_results/rawdataplot_AVG_humerus_radius.pdf")
op<-par(mar=c(5.1, 4.1, 0.2, 2.1))
plot(phylo_dat$AVG.humerus,phylo_dat$AVG.radius,type="p",col="red",xlab="AVG.humerus",ylab="AVG.radius",
     cex.lab=1.5,cex.axis=1.5)
par(op)
dev.off()
#--------------------------------------------------------------------------------------------
# Generating copula plot btw AVG.humerus and AVG.radius from contrasts.csv
pdf("./Results/Anolis_results/contrasts_csv_results/copula_AVG_humerus_radius.pdf")
v_contrasts_67<-phylo_dat_fn(data=phylo_dat,i=6,j=7,ploton = T)
dev.off()
```

\begin{figure}[!h]
\textbf{ \hspace{5.5 cm} (A) \hspace{7 cm} (B)  } \\
\begin{center}
\includegraphics[width=3 in]{./Results/Anolis_results/contrasts_csv_results/rawdataplot_AVG_humerus_radius.pdf}
\includegraphics[width=3 in]{./Results/Anolis_results/contrasts_csv_results/copula_AVG_humerus_radius.pdf}
\caption{\textbf{(A)} Plot of the raw AVG.radius vs. AVG.humerus phylogenetic traits. \textbf{(B)} Copula plot of AVG.radius vs. AVG.humerus phylogenetic traits. Obtained by applying ranks separately to each variable, and then dividing the ranks by $n+1$ where $n=`r nrow(v_contrasts_67)`$ is the number of contrasts (see Sec. \ref{Model_selection}). \label{fig_raw_cop_AVG_humerus_radius}}
\end{center}
\end{figure}

@@

We also tested statistically for asymmetry of tail dependence, which seems visible in Figure \ref{fig_raw_cop_soilCN}(B). 
Results (Table \ref{tab_soilCN_asym}) show that upper-tail dependence was, indeed, significantly stronger than lower-tail dependence in these data, answering the third part of question 1.

@@

 #atdres<-data.frame(Statistic=c("$\\cor_{0,0.1}-\\cor_{0.9,1}$",
 #                               "$\\Ps_{0,0.1}-\\Ps_{0.9,1}$",
 #                               "$\\Dsq_{0.9,1}-\\Dsq_{0,0.1}$"),
 atdres<-data.frame(Statistic=c("$\\cor_{first.slice}-\\cor_{last.slice}$",
                               "$\\Ps_{first.slice}-\\Ps_{last.slice}$",
                                "$\\Dsq_{first.slice}-\\Dsq_{last.slice}$"),
                   Kendall=rep("",3),
                   Spearman=rep("",3),stringsAsFactors = F)

@@

@@

The test of indepedence gave $p=$ `r BivMS_BMR$IndepTestRes` (to within the precision available from `BiCopIndTest` in the `VineCopula` package). We fitted several copulas
and got AIC values and upper- and lower-tail dependency estimates based on the fitted copulas (Table \ref{tab_BMRfit}). The table shows that Frank is much better supported (higher AIC and BIC weights) than the Gaussian copula, answering the first part of question 1 for these data.

@@

The model-averaged (AIC weights used for model averaging) lower-tail and upper-tail dependence statistic was $0$ as expected, because the contributions came from Frank and SBB8 which are both symmetric copula.
<!--$`r BivMS_BMR$relLTdep_AICw`$ and the model-averaged upper-tail dependence statistic was 
$`r BivMS_BMR$relUTdep_AICw`$. These values are distinct from what a Gaussian copula would give (i.e., 0), answering the second part of question 1 for these data. Furthermore, the difference of these two quantities (lower minus upper), 
$`r BivMS_BMR$relLTdep_AICw-BivMS_BMR$relUTdep_AICw`$, was different from 0, answering the third part of question 1. 
-->
We also tested that lowest-AIC copula Frank was a good-enough fit, giving $p$-value 
$`r BivMS_BMR$GofRes_CvM`$ (to within the precision available from `BiCopGofTest` in the `VineCopula` package) in the Cramer-von Mises-based test of the goodness of its fit, and $p$-value $`r BivMS_BMR$GofRes_KS`$ in the Kolmogorov-Smirnov-based test of the goodness of fit. The next section looks at tail dependence in a non-parametric way.

@@

```{r stat_BMR_graphic_approach, echo=F, results="hide", cache=T, cache.extra=list(seed,v_BMR,mtime("bivfunctionplot.R"),mtime("CopulaFunctions_flexible.R"),mtime("copsurrog2d.R"))}
source("bivfunctionplot.R")
set.seed(seed) 
# Now do NPA for v_BMR
# If "F"/"SBB8" is the best fit : then it should be symmetric tail dep (0) : Then still I need to do NPA ?
# May be : because it will also tell me how they are different from "N" copula in different sections?
source("bivfunctionplot.R")
stat_BMR_ga<-bivfunctionplot(v=v_BMR,resloc="./Results/BMR_results/",nametag="BMR_numbin_10",numbin = 10)
stat_BMR_ga_numbin5<-bivfunctionplot(v=v_BMR,resloc="./Results/BMR_results/",nametag="BMR_numbin_5",numbin = 5)
```

![Nonparametric tests for tail dependence and other deviations from a normal copula for BMR data. As described in Methods, each statistic was computed on the data using the ranges $0-0.1$, $0.1-0.2$, ..., $0.9-1$ (red lines), and was also computed using the same ranges for 1000 normal-copula surrogate datasets (dashed lines show 0.025 and 0.975 quantiles of these results and dotted lines show 0.005 and 0.995 quantiles). \label{fig_BMR_nonparam}](./Results/BMR_results/BMR_numbin_10_bivfunctionplot.pdf)

![Nonparametric tests for tail dependence and other deviations from a normal copula for BMR data. As described in Methods, each statistic was computed on the data using the ranges $0-0.2$, $0.2-0.4$, ..., $0.8-1$ (red lines), and was also computed using the same ranges for 1000 normal-copula surrogate datasets (dashed lines show 0.025 and 0.975 quantiles of these results and dotted lines show 0.005 and 0.995 quantiles). \label{fig_BMR_nonparam5}](./Results/BMR_results/BMR_numbin_5_bivfunctionplot.pdf)

```{r tab_BMR_asym, echo=F, results="markup"}
tab_BMR_asym<-make_table_stat_fn(data=stat_BMR_ga)
tab_BMR_asym_numbin5<-make_table_stat_fn(data=stat_BMR_ga_numbin5)
knitr::kable(tab_BMR_asym,
             format="pandoc",
             caption = "Results for tests of whether asymmetry of tail dependence was significant compared to a normal-copula null hypothesis (best fit is here Frank with symmetric tail dep)?, numbin=10 \\label{tab_BMR_asym}",
             booktabs=T)

knitr::kable(tab_BMR_asym_numbin5,
             format="pandoc",
             caption = "Results for tests of whether asymmetry of tail dependence was significant compared to a normal-copula null hypothesis (best fit is here Frank with symmetric tail dep)?, numbin=5 \\label{tab_BMR_asym_numbin5}",
             booktabs=T)
```

## Results: contrasts.csv data \label{Ex_contrasts}

### Model selection approach \label{Ex_contrasts_Model_selection}

```{r bivMS_contrasts_67, echo=F, cache=T, cache.extra=list(seed,v_contrasts_67,mtime("OurBiCopSelect.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
set.seed(seed)
# Now do Model Selection for [i,j]=[6,7]
# 6 : AVG. humerus
# 7 : AVG. radius
source("OurBiCopSelect.R")
families<-c(1,3:10,13:14,16:20)
BivMS_contrasts_67<-OurBiCopSelect(u1=v_contrasts_67[,1],u2=v_contrasts_67[,2],families=families,
                          level=0.05,AICBIC="AIC",numBSsmall=100,pthresh=0.2,numBSlarge=1000,
                          gofnormal=F,status=F)
saveRDS(BivMS_contrasts_67,file="./Results/Anolis_results/contrasts_csv_results/BivMS_contrasts_67.RDS")
```

The test of indepedence gave $p=$ `r BivMS_contrasts_67$IndepTestRes` (to within the precision available 
from `BiCopIndTest` in the `VineCopula` package). We fitted several copulas
and got AIC values and upper- and lower-tail dependency estimates based on the fitted copulas (Table \ref{tab_contrasts_67fit}). The table shows that several copulas are much better supported (higher AIC and BIC weights) than the Gaussian copula, answering the first part of question 1 for these data.

<!--Display model selection results for contrasts data in a table-->
```{r tab_contrasts_67fit, echo=F, results='markup'}
h<-BivMS_contrasts_67$InfCritRes[,c("copname","AICw","BICw","LTdep","UTdep")]
knitr::kable(h,
             format='pandoc',
             caption = "Fitting info for contrasts between AVG.humerus and AVG.radius dataset. \\label{tab_contrasts_67fit}",
             booktabs=T)
```

The model-averaged (AIC weights used for model averaging) lower-tail dependence statistic was `r BivMS_contrasts_67$relLTdep_AICw` and the model-averaged upper-tail dependence statistic was `r BivMS_contrasts_67$relUTdep_AICw`. These values are distinct from what a Gaussian copula would give (i.e., 0), answering the second part of question 1 for these data. Furthermore, the difference of these two quantities (lower minus upper), `r BivMS_contrasts_67$relLTdep_AICw-BivMS_contrasts_67$relUTdep_AICw`, was different from 0, answering the third part of question 1.

These results do convincingly show non-Gaussian copula structure, giving $p$-value `r BivMS_contrasts_67$GofRes_CvM` (to within the precision available from `BiCopGofTest` in the `VineCopula` package) in the Cramer-von Mises-based test of the goodness of its fit, and $p$-value `r BivMS_contrasts_67$GofRes_KS` in the Kolmogorov-Smirnov-based test of the goodness of fit. The next section looks at tail dependence in a non-parametric way.

### Non-parametric analysis of tail dependence \label{Ex_contrasts_67_NPA}

```{r stat_contrasts_67_graphic_approach, echo=F, results="hide", cache=T, cache.extra=list(seed,v_contrasts_67,mtime("bivfunctionplot.R"),mtime("CopulaFunctions_flexible.R"),mtime("copsurrog2d.R"))}
source("bivfunctionplot.R")
set.seed(seed) 
# Now do NPA for v_contrasts_67
source("bivfunctionplot.R")
stat_contrasts_67_ga<-bivfunctionplot(v=v_contrasts_67,resloc="./Results/Anolis_results/contrasts_csv_results/",
                                      nametag="contrasts_67",numbin = 4)
```

![Nonparametric tests for tail dependence and other deviations from a normal copula for contrasts_67 data. As described in Methods, each statistic was computed on the data using the ranges $0-0.1$, $0.1-0.2$, ..., $0.9-1$ (red lines), and was also computed using the same ranges for 1000 normal-copula surrogate datasets (dashed lines show 0.025 and 0.975 quantiles of these results and dotted lines show 0.005 and 0.995 quantiles). \label{fig_BMR_nonparam}](./Results/Anolis_results/contrasts_csv_results/contrasts_67_bivfunctionplot.pdf)

```{r tab_contrasts_67_asym, echo=F, results="markup"}
tab_contrasts_67_asym<-make_table_stat_fn(data=stat_contrasts_67_ga)
knitr::kable(tab_contrasts_67_asym,
             format="pandoc",
             caption = "Results for tests of whether asymmetry of tail dependence was significant compared to a normal-copula null hypothesis \\label{tab_contrasts_67_asym}",
             booktabs=T)
```
